---
title: "README"
author: "Zachary McCaw"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: TRUE
--- 

# Multiple Outcome Restricted Mean Survival Time

Zachary R. McCaw <br>
Updated: 2023-10-06

```{r global_options, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

### Overview

Suppose that, for each patient, $K$ time-to-event outcomes $(T_{1}, \dots, T_{K})$ outcomes are available, each potentially subject to censoring. Let $\delta_{k} = 1$ if the $k$th event is observed, and $\delta_{k} = 0$ if the $k$th event is censored. For a specified time window $(0, \tau)$, define the parameter:

$$
\theta = \sum_{k=1}^{K}\int_{0}^{\tau}S_{k}(t)dt,
$$
where $S_{k}(t) = \mathbb{P}(T_{k} > t)$. $\theta$ can be interpreted as the cumulative area under the curve (AUC) for the $K$ survival functions $(S_{1}, \dots, S_{K})$. This package performance inference on the difference and ratio of $\theta$, comparing two independent treatment arms.


### Data

Example data can be simulated using the `GenData` function:
```{r}
library(MRMST)
data <- MRMST::GenData(n_subj = 100, n_event = 2)
head(data)
```

### One-Sample Problem

To estimate the AUC for a single treatment arm in the case of $K=2$ time-to-event outcomes:

```{r}
# Single arm estimation.
one_sample <- MRMST::OneSample(
  statuses = data[, c("status1", "status2")],
  times = data[, c("time1", "time2")]
)
show(one_sample)
```

### Two-Sample Problem

To perform inference on the AUC comparing two treatment arms:

```{r}

# Simulate data with a treatment effect.
arm <- rep(c(0, 1), each = 50)
data <- MRMST::GenData(
  covariates = data.frame(arm = arm), 
  beta_event = 1.0,
  n_event = 2
)

# Comparison of two arms.
two_sample <- MRMST::TwoSample(
  arm = data$arm,
  statuses = data[, c("status1", "status2")],
  times = data[, c("time1", "time2")]
)
show(two_sample)
```






